---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# westerlund

<!-- badges: start -->
<!-- badges: end -->

`westerlund` is a toolkit for single-cell RNA-sequencing analysis built on top of Seurat providing functions to quickly and comprehensively assess clustering results. Inspired by many publications and prior packages (highlighted below), `westerlund` enables data sketching, cluster bootstrapping, automatic and assisted resolution scanning, and metric visualizations. 

## Why `westerlund`?
The functions within `westerlund` provide additional critical functionality for common single-cell analyses that have been implemented in a range of outstanding publications, including:   
1. Supporting functions based on the recommendations by the authors of [pipeComp](https://doi.org/10.1186/s13059-020-02136-7),  
1. Leiden clustering implemented in the `leidenbase` algorithm by the [Trapnell Lab](https://github.com/cole-trapnell-lab/leidenbase),  
1. Interpretable resolution and k-nearest neighbors scanning for determining the optimal cluster granularity utilizing the `future` framework,  
1. Heuristic metrics to determine appropriate PCs to utilize for further dimensionality reduction and clustering,  
1. Cluster robustness measures including SVM classification and adjusted Rand index random seed and subset sampling,  
1. Functions to quickly assess cluster significance based on DE genes and merge clusters based on size and/or user flagging,  
1. Gene list scoring using scaled expression instead of normalized expression as utilized by [Smillie et al.](https://doi.org/10.1016/j.cell.2019.06.029),  
1. Bayesian classifiers as implemented by [Zeymour et al.](https://doi.org/10.1038/s41590-018-0051-0),  
1. Gene specificity index calculations as proposed by [Tosches et al.](https://doi.org/10.1126/science.aar4237),  
1. Max-to-second expression ratios as utilized in [Zilions et al.](https://doi.org/10.1016/j.immuni.2019.03.009).

These functions are plug and play with `Seurat` objects and can enable easy expansion of available scRNA-seq analytical tools.

## Installation

```
# install.packages("devtools")
devtools::install_github("joshpeters/westerlund")
```
## Examples

```{r echo = T, eval = F}
library(tidyverse)
library(Seurat)
library(SeuratData)
library(westerlund)

#SeuratData::InstallData("pbmc3k")
#SeuratData::InstalledData()
# object <- SeuratData::LoadData("pbmc3k")
# object$seurat_annotations <- as.character(object$seurat_annotations)
# object$seurat_annotations[is.na(object$seurat_annotations)] <- "Unassigned"
# Idents(object) <- "seurat_annotations"
# object <- ReduceDims(seurat.obj = object, n.var.features = 3000, n.pcs = 50, remove.genes = NULL)
# object <- Seurat::RunUMAP(object = object, reduction = "pca", reduction.name = "pca_umap",
#                   dims = 1:object@misc$pca_metrics$tp_pc, seed.use = 1, verbose = FALSE)
# plot <- DimPlot(object, reduction = "pca_umap", group.by = "seurat_annotations", pt.size = 1, shuffle = TRUE, label = FALSE, label.size = 6, repel = TRUE)
# plot + theme_classic(base_size = 18) +
#   #NoLegend() +
#   RemoveAxes() + RemoveBackgrounds(outline = FALSE) + SpaceAxesTitles() +
#   colorspace::scale_color_discrete_qualitative("Dark 3") +
#   labs(x = "UMAP1", y = "UMAP2", title = "")
# markers <- DefineMarkers(object, "seurat_annotations")
# Bcell_markers <- markers %>% filter(group == "B") %>% top_n(100, auc) %>% pull(feature)
# object <- AddModuleScoreScaled(object, list(Bcell_markers), name = "SMS")
# object@meta.data[, grep("^MS", colnames(object[[]]))] <- NULL
# object <- AddModuleScore(object, list(Bcell_markers), name = "MS", nbin = 20, ctrl = 100)


```
